---
title: "Relationships between employment sectors"
author: "David J. Harris"
date: "10/31/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The US Census Bureau collects data every year about economic activity across
sectors of the US economy and across regions as part of the County Business 
Patterns (CBP) program. The CBP data for 2014 is available
from their website, where it can be downloaded and imported easily:

```{r}
library(tidyverse)
library(readr)
```


```{r, cache = TRUE}
if(!file.exists("CB1400A11.zip")){
  download.file("http://www2.census.gov/econ2014/CB/sector00/CB1400A11.zip",
                "CB1400A11.zip")
  unzip("CB1400A11.zip")
}
raw_CB_data = read_delim("CB1400A11.dat", delim = "|")
```


For this analysis, I will focus on the county level (i.e. regions whose
value for `COUNTY` is not 1 or 999), and on the three-digit version of the
North American Industry Classification System (NAICS2012), which provides
enough granularity to get a good picture of a county's economy, but which
has a manageable number of categories.

The following code generates a table with one row per county and one
column per employment sector, with entries describing the number of 
employees working in each sector in the first quarter of 2014.

```{r}
spread_data = raw_CB_data %>% 
  filter(nchar(as.character(NAICS2012)) == 3) %>% 
  filter(COUNTY %in% 2:998) %>% 
  distinct(GEO_ID, GEO_TTL, NAICS2012_TTL, EMP) %>% 
  spread(NAICS2012_TTL, EMP, fill = 0)


counts = select(spread_data, -starts_with("GEO")) %>% 
  filter(rowSums(.) > 0) %>% 
  select(which(colSums(.) > 0))
```


We can visualize the correlations among percent employment in each sector
as follows:

```{r, fig.height=1000, fig.width=1000}
reshape2::melt(cor(as.matrix(counts) / rowSums(counts)) * upper.tri(cor(counts))) %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() + 
  scale_fill_gradient2() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Surprisingly, the correlations are generally quite weak (96% below
0.1), indicating that the sectors tend not to rely heavily on one another
at the county level. 

The few highly-correlated pairs are generally unsurprising, however. For
example, management employment is correlated with financial activity and
motion picture production, which typically involve large companies, and tends
to be a smaller sector of the economy in regions dominated by gas stations
and restaurants.

